The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-08 rational.tex
particles

%%%%%%%%%%%%%%%%%%%%
\subsection{Syntax}

@@

The syntactic notation described above is often obscure. Following the
tradition established for monoidal
categories~\cite{springerlink:10.1007/978-3-642-12821-94}, we present a
graphical language that conveys the intuitive semantics of the language
(which is formalized in the next section). The general idea of the graphical
notation is that combinators are modeled by ``wiring diagrams'' or
``circuits''; values are modeled as ``particles'' that may appear on the
wires; and evaluation is modeled by the flow of particles along the wires. We
now proceed to give the graphical notation for each combinator:

\begin{itemize}
\item The simplest sort of diagram is the {{id : b <-> b}} combinator which
  is simply represented as a wire labeled by its type {{b}}. In more complex
  diagrams, if the type of a wire is obvious from the context, it may be
  omitted.
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/b-wire.pdf]
\includegraphics{diagrams/thesis/b-wire.pdf}
}
\end{center}

@@

@@

@@

\item Identity morphisms and their duals ({{uniti, unite, zeroi}} and
  {{zeroe}}) are represented as follows:
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr1.pdf]
\includegraphics{diagrams/thesis/uniti.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl1.pdf]
\includegraphics{diagrams/thesis/unite.pdf}
}
\end{center}  
\end{multicols}
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr0.pdf]
\includegraphics{diagrams/thesis/zeroi.pdf}
}
\end{center}
\columnbreak
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl0.pdf]
\includegraphics{diagrams/thesis/zeroe.pdf}
}
\end{center}
\end{multicols}
Since the monoidal units can be freely introduced and eliminated, in
many diagrams they are omitted and dealt with explicitly when they
are of special interest.

>>>>

%%%%%%%%%%%%%%%%%%%%
\subsection{Syntax} 
\label{sec:pi-syntax}

@@

The syntactic notation described in Sec \ref{sec:pi-syntax} is often
obscure and hard to read.  Following the tradition established for
monoidal categories~\cite{springerlink:10.1007/978-3-642-12821-94}, we
present a graphical language that conveys the intuitive semantics of
the language (which is formalized in the next section).  For the rest
of the paper {{langRev}} programs will be presented graphically.

The general idea of the graphical notation is that combinators are
modeled by ``wiring diagrams'' or ``circuits''; values are modeled as
``particles'' that may appear on the wires; and evaluation is modeled
by the flow of particles along the wires.

%% We now proceed to give the graphical notation for each combinator:

\begin{itemize}
\item The simplest sort of diagram is the {{id : b <-> b}} combinator
  which is simply represented as a wire labeled by its type {{b}}, as
  shown on the left. In more complex diagrams, if the type of a wire
  is obvious from the context, it may be omitted.

  \begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/b-wire.pdf]
\includegraphics{diagrams/thesis/b-wire.pdf}
}
\end{center}

DIAGRAM    
  \end{multicols}

When tracing a computation, one might imagine a value {{v}} of of type
{{b}} on the wire, as shown on the right.

@@

In the case of products represented by a pair of wires, when tracing
execution using particles, one should think of one value/particle on
each wire.

@@

When tracing the execution of two additive wires, a value can reside
on only one of the two wires.

@@

The crisscrossing wires have the right semantics in that a value on
the top wire, ends up on the bottom wire and vice versa.

\item Identity morphisms and their duals ({{uniti, unite, zeroi}} and
  {{zeroe}}) are represented as follows:
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr1.pdf]
\includegraphics{diagrams/thesis/uniti.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl1.pdf]
\includegraphics{diagrams/thesis/unite.pdf}
}
\end{center}  
\end{multicols}
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr0.pdf]
\includegraphics{diagrams/thesis/zeroi.pdf}
}
\end{center}
\columnbreak
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl0.pdf]
\includegraphics{diagrams/thesis/zeroe.pdf}
}
\end{center}
\end{multicols}
Note that since there is no value of type 0, there can be no particle
on a wire of type {{0}}.  Since the monoidal units can be freely
introduced and eliminated, in many diagrams they are omitted and dealt
with explicitly when they are of special interest.
